---
title: "HW 03 - Great Lakes Fish Datasets"
author: "Kristina Totoraitis"
date: "02.23.23"
output: html_document
---

## Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse)
```

```{r load-data, message = FALSE}
catch_data <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-06-08/fishing.csv')

fish_stock <- read_csv("data/great-lakes-fish-stock.csv")
```

## Exercises

### Exercise 1

```{r catch-data-subset}
catch_data <- catch_data %>%
  select(year, lake, species, region, values)
```
```{r fish-stock-subset}
fish_stock <- fish_stock %>%
  select(YEAR, MONTH, LAKE, STATE_PROV, SPECIES, WEIGHT)

```

### Exercise 2

```{r}
glimpse(catch_data)
glimpse(fish_stock)

```

### Exercise 3
```{r filter-catch, eval = TRUE}
catch_data <- catch_data %>% 
  filter(region == "U.S. Total",
         species == "Chinook Salmon",
         lake == "Michigan")
```
```{r filter-stock, eval = TRUE}
fish_stock <- fish_stock %>% 
  filter(LAKE %in% c("MI", "WI", "IL", "IN"),
         SPECIES == "CHS",
         LAKE == "MI")
```


### Exercise 4
There are `r nrow(catch_data)` columns in the catch data set.
The min year is `r min(catch_data$year)` for the catch data set.
The max year is `r max(catch_data$year)` for the catch data set.
There are `r nrow(fish_stock)` columns in the fish stock data set.
There min year is `r min(fish_stock$YEAR)` for the fish stock data set. 
The max year is `r max(fish_stock$YEAR)` for the fish stock data set. 


### Exercise 5
```{r eval=TRUE}
fish_stock_yearly<- fish_stock%>%
  group_by(YEAR)%>%
summarize( STOCK_WEIGHT_YR= sum(WEIGHT, na.rm=TRUE))

```

```{r finalize-fish-stock, eval = TRUE}
fish_stock_yearly <- fish_stock_yearly %>% 
  rename(year = YEAR) %>% 
  mutate(stock_weight_yr_lbs = as.integer(STOCK_WEIGHT_YR * 0.454 / 1000)) %>% # 1 lb = 1 kg * 0.454
  select(year, stock_weight_yr_lbs)
```



```{r finalize-catch-data, eval = FALSE}
catch_data <- catch_data %>% 
  rename(catch_weight_yr_lbs = values)
```

### Exercise 6

```{r left-join}
catch_data%>%
  left_join( fish_stock_yearly)
```

Add narrative here.

```{r right-join}
right_join ( catch_data, fish_stock_yearly)
```

Add narrative here.

```{r inner-join}
inner_join(catch_data, fish_stock_yearly)
```

Add narrative here.

```{r full-join}
full_join(catch_data, fish_stock_yearly)
```
```{r anti-join-v1, eval = TRUE}
catch_data %>% 
  anti_join(fish_stock_yearly)
```

```{r anti-join-v2, eval = FALSE}
fish_stock_yearly %>% 
  anti_join(catch_data)
```



### Exercise 7
```{r pivot-longer, eval = FALSE}
catch_data %>% 
  left_join(fish_stock_yearly) %>% 
  pivot_longer(names_to = "activity",
               values_to = "weight_lbs",
               cols = 5:6)
```


### Exercise 8


```{r final-visualization, eval = FALSE}
catch_data %>% 
  left_join(fish_stock_yearly) %>% 
  pivot_longer(names_to = "activity",
               values_to = "weight_lbs",
               cols = 5:6)%>%
  ggplot(aes( x=year, y= weight_lbs, color= activity))+ geom_line()+ theme_bw()
```